Marlowe: financial contracts on blockchain

### 2018-05-18 tex/paper.tex
Fix various details and typos derived from review

First, it will need to observe different kinds of varying quantities including, for example, the current time, the 
current block number and, random numbers, as well as ``real world'' quantities like ``the price of oil'' or ``the 
exchange rate between currencies A and B''. 
As the examples illustrate, observables come both from aspects of the blockchain (e.g.\ the current block number) and 
externally. In the latter case it will be necessary to agree a trusted oracle or beacon giving the value.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themsleves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of payments by 
a participant. <oreover it is also possible for a participant to input an arbitrary value (which we term a ``choice''). 
The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

While informally we might see a commitment to something as being indefinite, it is important to realise that on 
blockchain a commitment needs to have a timeout so that progress can be forced in a contract. After the timeout period the 
cash can be refunded through the user creating a transaction to reclaim the cash.
Information about the commitments currently in force forms the \haskellinline{State}, which can be  modified 
at each execution step.

@@

\begin{itemize}
\item It is assumed that cryptographic functions and operations are provided by a layer external to the system, and so 
they need not be modelled explicitly.
\item We assume that time is ``coarse grained'' and measured by block number, so that in particular timeouts are 
delimited using block numbers. 
\item 
Making a commitment is not something that a contract can perform; rather, it 
can request that a commitment is made, but that then has to be established externally: hence the input of (a set of) 
commitments at each step.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of 
the \haskellinline{stepBlock} function below).
\end{itemize}

@@

In addition to 
calling \haskellinline{stepAll}, \haskellinline{stepBlock} will first enable expired cash commitments to be refunded and record any choices made at that step. 
The functions \haskellinline{stepAll} and \haskellinline{stepBlock} have the same type as \haskellinline{step} itself.

@@

The type of contracts is given by the Haskell data type

\begin{haskellcode}
data Contract =
   Null |
   CommitCash IdentCC Person Money Timeout Timeout Contract Contract |  
   RedeemCC IdentCC Contract |
   Pay IdentPay Person Person Money Timeout Contract |  
   Both Contract Contract |
   Choice Observation Contract Contract |
   When Observation Timeout Contract Contract   
   \end{haskellcode}
Informally, this type provides a \haskellinline{Null} contract, which does nothing. The next three constructs form contracts that do something, and then continue according to another contract (which is one of the components of the original contract). \haskellinline{CommitCash} will wait for a participant to make a commitment,  
\haskellinline{RedeemCC} allows for a commitment to be redeemed, and \haskellinline{Pay} for a payment to be made to another participant.

@@

What many of the contracts also have are timeouts that also determine their behaviour. 
\subsection{The \haskellinline{step} function}

In this section we explain the detailed behaviour of \haskellinline{Contracts} through a discussion of the \haskellinline{step} function.

@@

\smallskip
\noindent
\haskellinline{CommitCash ident person val start_timeout end_timeout con1 con2}  For this contract to make 
progress,\begin{itemize}
\item either before the timeout \haskellinline{start_timeout} the user \haskellinline{person} makes a cash commitment of 
value 
\haskellinline{val} and timeout \haskellinline{end_timeout}  with the identifier \haskellinline{ident}, 
\item
or the timeout \haskellinline{start_timeout} is exceed:
\end{itemize}

@@

If the cash is committed successfully and the timeout \haskellinline{end_timeout} is reached, then it is impossible to 
further spend the committed cash, and any unspent funds can be reclaimed by participant \haskellinline{person}.
 This is performed by the \haskellinline{stepBlock} function, as noted above.

\medskip
\noindent
\haskellinline{RedeemCC ident con} (\haskellinline{CC} stands for cash commitment.) For this contract to make 
progress, the creator of the cash commitment with identifier \haskellinline{ident} is required to redeem the unspent 
funds in that commitment; the contract then continues as \haskellinline{con}, and the action 
\haskellinline{CommitRedeemed} is produced.

@@

\begin{haskellcode}
step inp st c@(Pay idpay from to val expi con) os
  | expired (blockNumber os) expi = (st, con, [ExpiredPay idpay from to cval])
  | right_claim =
    if committed st from bn >= cval
      then (newstate, con, [SuccessfulPay idpay from to cval])
      else (st, con, [FailedPay idpay from to cval])
  | otherwise = (st, c, [])
  where
    cval = evalMoney st val
    newstate = stateUpdate st from to bn cval
    bn = blockNumber os
    right_claim =
      case Map.lookup (idpay, to) (rp inp) of
        Just claimed_val -> claimed_val == cval
        Nothing -> False
\end{haskellcode}
By `available' we mean that sufficient commitments have been made and not yet expired to cover the payment; in this 
case the payment uses the currency allocated by the cash commitments made by \haskellinline{from} that expire the 
earliest.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
And an identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} to 
be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either participant \haskellinline{1} deserves a refund, which we 
represent with the number $0$; or participant \haskellinline{2} deserves a payment, which we represent with the number 
$1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as their 
participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then we can easily write a function that returns an observation that is true if and only if at least one of the 
participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{c} as follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is true if and only if at least 
two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed in 
choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose 1 2 3 0)
                                 (two_chose 1 2 3 1))
                          90
                          (Choice (two_chose 1 2 3 1)
                                  (Pay iP1 1 2 (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

The outermost primitive \haskellinline{CommitCash} allows the participant \haskellinline{1} to commit 450 ADA before 
block 10, with the promise that money will be released on block 100 if they are not claimed before that.

The next primitive, \haskellinline{When}, waits for one of three things to happen:
\begin{enumerate}
 \item The observation became true because two out of three have chosen $0$. 
 \item The observation became true because two out of three have chosen $1$.
 \item The observation remained false until 90 was published in the blockchain.
\end{enumerate}

If the third option happens, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to participant \haskellinline{1} unless two out of three chose option $1$; in the later 
case, \haskellinline{Pay} will give participant \haskellinline{2} the opportunity to claim the funds available in 
the commitment with identifier \haskellinline{iCC1} before block 100 (by using the identifier \haskellinline{iP1} for 
the claim). If the funds are not claimed before block 100 they will also be refunded to 
participant \haskellinline{1}.

@@

Blockly's editor allows the user to visualise and edit
smart-contracts as interlocking blocks that can be dragged and dropped
like pieces of a jigsaw puzzle. % (see Figure~\ref{fig:detail-of-block}).
Meadow also provides functionality to generate syntactically correct
and formatted code (that is displayed in the upper right corner of the application),
and to convert the code back to its block representation.

@@

@@

Commitments and values of observables need to be recorded somehow, and this raises the potential issue about bloating 
the chain with data. In general, it will only be necessary to store a signed hash value of the observable, and this will 
keep data usage bounded.

@@

The Findel project~\cite{findel} examines financial contracts on the Ethereum platform, based on the 
seminal~\cite{PeytonJones:2000}, and the authors note that payments need to be bounded; this is made concrete in our 
account by our notion of commitments. They take no account of commitments or timeouts as we do below, but it should be 
noted that the Ethereum platform is more powerful than the one that we target in this paper.

@@

In this paper we have presented Marlowe, a DSL for financial contracts on blockchains, based on earlier work on contracts~\cite{PeytonJones:2000}, together with examples of its use. We have seen that to make this operational on blockchain we need to add commitments and timeouts, and to design a semantics that reflects these. As we saw in Section~\ref{sec:abstraction}, Marlowe has been designed to make as few demands as possible on the underlying blockchain: it can be implemented on UTxO or account-based blockchains, for example.

>>>>

First, it will need to observe different kinds of varying quantities including, for example, the current time, the 
current block number and, random numbers, as well as ``real world'' quantities like ``the price of oil'' or ``the 
exchange rate between currencies A and B''. 
As the examples illustrate, observables come both from aspects of the blockchain (e.g.\ the current block number) and 
externally. In the latter case, it will be necessary to agree a trusted oracle or beacon giving the value.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themsleves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of payments by 
a participant. Moreover it is also possible for a participant to input an arbitrary value (which we term a ``choice''). 
The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

While informally we might see a commitment to something as being indefinite, it is important to realise that, on
blockchain, a commitment needs to have a timeout so that progress can be forced in a contract. After the timeout period the 
cash can be refunded through the user creating a transaction to reclaim the cash.
Information about the commitments currently in force forms the \haskellinline{State}, which can be  modified 
at each execution step.

@@

\begin{itemize}
\item It is assumed that cryptographic functions and operations are provided by a layer external to the system, and so 
they need not be modelled explicitly.
\item We assume that time is ``coarse grained'' and measured by block number, so that, in particular, timeouts are 
delimited using block numbers. 
\item 
Making a commitment is not something that a contract can perform; rather, it 
can request that a commitment is made, but that then has to be established externally: hence the input of (a set of) 
commitments at each step.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of 
the \haskellinline{stepBlock} function below).
\end{itemize}

@@

In addition to 
calling \haskellinline{stepAll}, \haskellinline{stepBlock} will first enable expired cash commitments to be refunded and record, in the state, any choices made at that step. 
The functions \haskellinline{stepAll} and \haskellinline{stepBlock} have the same type as \haskellinline{step} itself.

@@

The type of contracts is given by the following Haskell data type:

\begin{haskellcode}
data Contract =
   Null |
   CommitCash IdentCC Person Money Timeout Timeout Contract Contract |  
   RedeemCC IdentCC Contract |
   Pay IdentPay Person Person Money Timeout Contract |  
   Both Contract Contract |
   Choice Observation Contract Contract |
   When Observation Timeout Contract Contract   
   \end{haskellcode}
Informally, this type provides a \haskellinline{Null} contract, which does nothing. The next three constructs form contracts that do something, and then continue according to another contract (which is one of the components of the original contract). \haskellinline{CommitCash} will wait for a participant to make a commitment,  
\haskellinline{RedeemCC} allows for a commitment to be redeemed, and \haskellinline{Pay} for a payment between participants to be claimed by the recipient.

@@

Additionally, many of the contracts have timeouts that also determine their behaviour. 
\subsection{The \haskellinline{step} function}

In this section, we explain the detailed behaviour of \haskellinline{Contracts} through a discussion of the \haskellinline{step} function.

@@

\smallskip
\noindent
\haskellinline{CommitCash ident person val start_timeout end_timeout con1 con2}  For this contract to make 
progress,\begin{itemize}
\item either, before the timeout \haskellinline{start_timeout}, the user \haskellinline{person} makes a cash commitment of 
value 
\haskellinline{val} and timeout \haskellinline{end_timeout}  with the identifier \haskellinline{ident}, 
\item
or the timeout \haskellinline{start_timeout} is exceed:
\end{itemize}

@@

If the cash is committed successfully and the timeout \haskellinline{end_timeout} is reached, then it is impossible to 
further spend the committed cash, and any unspent funds can be reclaimed by participant \haskellinline{person}.
 This is enforced by the \haskellinline{stepBlock} function, as noted above.

\medskip
\noindent
\haskellinline{RedeemCC ident con} (\haskellinline{CC} stands for cash commitment.) For this contract to make 
progress, the creator of the cash commitment with identifier \haskellinline{ident} is allowed to redeem the unspent 
funds in that commitment; the contract then continues as \haskellinline{con}, and the action 
\haskellinline{CommitRedeemed} is produced.

@@

\begin{haskellcode}
step inp st c@(Pay idpay from to val expi con) os
  | expired (blockNumber os) expi = (st, con, [ExpiredPay idpay from to cval])
  | right_claim =
    if committed st from bn >= cval
      then (newstate, con, [SuccessfulPay idpay from to cval])
      else (st, con, [FailedPay idpay from to cval])
  | otherwise = (st, c, [])
  where
    cval = evalMoney st val
    newstate = stateUpdate st from to bn cval
    bn = blockNumber os
    right_claim =
      case Map.lookup (idpay, to) (rp inp) of
        Just claimed_val -> claimed_val == cval
        Nothing -> False
\end{haskellcode}
By `available' we mean that sufficient commitments have been made and not yet expired to cover the payment; in this 
case, the payment uses the currency allocated by the cash commitments made by \haskellinline{from} that expire the 
earliest.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
And an identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} to 
be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either participant \haskellinline{1} deserves a refund, which we 
represent with the number $0$; or participant \haskellinline{2} deserves a payment, which we represent with the number 
$1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as their 
participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then, we can easily write a function that returns an observation that is \haskellinline{True} if and only if at least one of the 
participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{val} as follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is \haskellinline{True} if and only if at least 
two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed in 
choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose 1 2 3 0)
                                 (two_chose 1 2 3 1))
                          90
                          (Choice (two_chose 1 2 3 1)
                                  (Pay iP1 1 2 (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

The outermost primitive \haskellinline{CommitCash} allows the participant \haskellinline{1} to commit \haskellinline{450} ADA before 
block \haskellinline{10}, with the promise that money will be released on block \haskellinline{100} if they are not claimed before that.

The next primitive, \haskellinline{When}, waits for one of three things to happen:
\begin{enumerate}
	\item The observation became \haskellinline{True} because two out of three have chosen \haskellinline{0}. 
	\item The observation became \haskellinline{True} because two out of three have chosen \haskellinline{1}.
	\item The observation remained \haskellinline{False} until \haskellinline{90} was published in the blockchain.
\end{enumerate}

If the third option happens, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to participant \haskellinline{1} unless two out of three chose option \haskellinline{1}; in the later 
case, \haskellinline{Pay} will give participant \haskellinline{2} the opportunity to claim the funds available in 
the commitment with identifier \haskellinline{iCC1} before block \haskellinline{100} (by using the identifier \haskellinline{iP1} for 
the claim). If the funds are not claimed before block \haskellinline{100} they will also be refunded to 
participant \haskellinline{1}.

@@

Blockly's editor allows the user to visualise and edit
smart-contracts as interlocking blocks that can be dragged and dropped
like pieces of a jigsaw puzzle. % (see Figure~\ref{fig:detail-of-block}).
Meadow also provides functionality to generate syntactically correct
and formatted code (that is displayed in the upper right corner of the application),
and to convert the code back to its blockly representation.

@@

Nevertheless, there already exists in Bitcoin a mechanism that allows the combination of several chained transactions within a single block while using the UTxO model, even though it is used to solve a different problem and it is not widely spread yet. This is the case of "child pays for parent", which allows the recipient of an unpublished transaction to issue another transaction that spends the former, with the aim of including enough fees for both transactions to have enough incentive to be included by a miner in a block (in which case both transactions would be included in the same block).

The same mechanism could be used for allowing miners to aggregate several transactions that act as a chain in the same block. The clients would only need to monitor the transaction pool in order to send transactions that can be chained with the existing ones.

@@

Inputs and values of observables need to be recorded somehow, and this raises the potential issue about bloating 
the chain with data. In general, it will only be necessary to store a signed hash value of the observable, and this will 
keep data usage bounded; but the full information may still need to be posted in case of dispute.

@@

The Findel project~\cite{findel} examines financial contracts on the Ethereum platform, based on the 
seminal~\cite{PeytonJones:2000}, and the authors note that payments need to be bounded; this is made concrete in our 
account by our notion of commitments. They take no account of commitments or timeouts as our approach does, but it should be 
noted that the Ethereum platform is more powerful than the one that we target in this paper.

@@

In this paper, we have presented Marlowe, a DSL for financial contracts on blockchains, based on earlier work on contracts~\cite{PeytonJones:2000}, together with examples of its use. We have seen that to make this operational on blockchain we need to add commitments and timeouts, and to design a semantics that reflects these. As we saw in Section~\ref{sec:abstraction}, Marlowe has been designed to make as few demands as possible on the underlying blockchain: it can be implemented on UTxO or account-based blockchains, for example.
